Bayesian model in JAGS

### 2023-01-27 4_RSV_disease_burden/4_RSV_disease_burden.Rmd
update

- Section 1: Background knowledge of estimating respiratory virus attributable disease burden (hospitalizations/mortality etc) with statistical models.
\
\
- Section 2: Background knowledge of Hierarchical Bayesian regression.
\
\
- Section 3: Background knowledge of using Hierarchical Bayesian regression to estimate RSV attributable hospitalizations.
\
\
- Section 4: R code for running the Hierarchical Bayesian regression to estimate RSV attributable hospitalizations in older adults by age and risk groups.

@@

## Section 2: Background knowledge of Hierarchical Bayesian regression

>>>>

- Section 1: Background knowledge of estimating respiratory virus attributable disease burden (hospitalizations/mortality etc) with statistical models.
\
\
- Section 2: Background knowledge of using Hierarchical Bayesian regression to estimate RSV attributable hospitalizations.
\
\
- Section 3: R code for running the Hierarchical Bayesian regression to estimate RSV attributable hospitalizations in older adults by age and risk groups.

@@

## Section 2: Background knowledge of using Hierarchical Bayesian regression to estimate RSV attributable hospitalizations

### Background knowledge of Hierarchical Bayesian regression

Hierarchical Bayesian regression can make population inference based on group and individual observations. The benefits of using a Hierarchical Bayesian structure is that it can shrink the uncertainty of inference while allow between-group variability.

## Background knowledge of Hierarchical Bayesian regression: example

For example, if we would like to estimate the baseline hospitalizations $\beta_{0jk}$ in age $j$ and socioeconomic group $k$, we can treat this parameter as (1) independent, meaning that the baseline hospitalizations are different in each group and we will need to estimate 27 parameters $\beta_{0j=1,k=1}$, $\beta_{0j=1,k=2}$...$\beta_{0j=9,k=3}$, (2) the same, meaning that the baseline hospitalizations are the same in all age and socioeconomic group and we will only need to estimate one parameter $\beta_0$, and (3) random draws from the a normal distribution, meaning that baseline hospitalizations are normally distributed in the whole population and baseline hospitalization of each subpopulation $jk$ is an realization of the distribution (a point in the distribution). In this case, we will estimate the mean and variations of the normal distribution and then get the subpopulation estimates.

## Further readings on Hierarchical Bayesian regression

- [Bayesian hierarchical modeling by wikipedia](https://en.wikipedia.org/wiki/Bayesian_hierarchical_modeling)

- [Chapter 10 Bayesian Hierarchical Modeling by Drs. Jim Albert and Jingchen Hu](https://bayesball.github.io/BOOK/bayesian-hierarchical-modeling.html)

- [Hierarchical Bayesian Modeling by Dr. Angie Wolfgang](https://astrostatistics.psu.edu/RLectures/hierarchical.pdf)
